---
title: "Data linkage"
author: "Coralie Williams"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    toc_collapsed: yes
    toc_depth: 3
    theme: cerulean
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Email from the 22/11/2022

------------------------------------------------------------------------

#### Problem 1

*In a nutshell, I am trying to create a new variable based on another one in my data frame. E.g. an ethnicity_group variable that is a value of 1 if the pre-existing variable Ethnicity is North African, 2 if North East Asian, 3 if North West European etc.*

------------------------------------------------------------------------

For demonstration, the following generates a dataframe of 90 observations/rows with 9 different types of 'fake' ethnicities:

```{r fakedata}
# Generate fake data (maded up ethnicities)
ethnicity_types <- c("NorthAfrican", "NorthEastAsian", "NorthWestEuropean",
               "SouthAfrican", "SouthEastAsian", "SouthWestEuropean",
               "WestAfrican", "WestAsian", "EastEuropean")

# Repeat the 9 ethnicities 10 times
ethnicity <- rep(ethnicity_types, 10)

# Save as a dataframe
dat <- as.data.frame(ethnicity)
```

First step, check and make sure the ethnicity variable is a factor:

```{r factor_var}
# Save ethnicity variable as a factor variable
dat$ethnicity <- as.factor(dat$ethnicity)
```

Derived the new variable `ethnicity_group` using the `as.numeric` function.

```{r ethn_group}
# Derived new variable "ethnicity_group" that represents each ethnicity group as a number
dat$ethnicity_group <- as.numeric(dat$ethnicity)
dat$ethnicity_group
```

Then obtain a summary output table of the correspondence between each ethnicity and number associated to it.

```{r}
# View the correspondence between each ethnicity and number
cbind(levels(dat$ethnicity), levels(as.factor(dat$ethnicity_group)))
```

------------------------------------------------------------------------

#### Problem 2

*Also trying to create a variable that inserts a date from one column in the data frame, depending on a value in another column. i.e. I have duplicate records because some people have 5 intervention dates. I want to create variables for intervention_date, intervention_date2, intervention_date3 etc. Managed to create an intervention_seq variable but no luck from there!*

------------------------------------------------------------------------

```{r, message=FALSE}
# Load packages to read csv file and view tables in a nice format
library(knitr)
library(readr)

# Read example dataset
exdat <- read_csv("data/example.csv")

### view output table
kable(exdat)
```

Use `pivot_wider` function from `tidyr` package to obtain table that has unique `PPN` variable values for each intervention date and type in columns:

-   `id_cols`: variable that uniquely identifies
-   `names_from`: variables that you use to get the columns
-   `values_from`: variables that will be used to get the cell values from

```{r, message=FALSE}
# Load tidyr package
library(tidyr)

# Obtain new table 
exdat_new <- pivot_wider(exdat, 
                         id_cols=c(PPN),
                         names_from=intervention_seq,
                         values_from=c(Intervention_Date, Intervention_Type))

### view output table
kable(exdat_new)
```

------------------------------------------------------------------------

### Email from the 28/11/2022

#### Change empty 'string' to missing value NA in R

After reading in dataset format variables so that all empty spaces are represented as NA's. Do this before changing any variable into a factor.

```{r, message=F}
# Generate fake data including empty strings
ethnicity_types <- c("NorthAfrican", "NorthEastAsian", "NorthWestEuropean",
               "SouthAfrican", "SouthEastAsian", "SouthWestEuropean",
               "WestAfrican", "WestAsian", "EastEuropean", "")

# repeat each character 10 times
ethnicity <- rep(ethnicity_types, 10)

# Save as a dataframe
dat <- as.data.frame(ethnicity)

# Replace empty strings into NA
library(dplyr)
dat <- dat %>% 
  mutate(ethnicity=ifelse(ethnicity=="",NA,ethnicity))

```

#### Appending additional variables to dataframe when using pivot_wider

If we want to add more variables that identify each observation this can be done by using the "id_cols" argument.

For example, if we would like to add "DOB" and "Gender" variables to the example above:

`exdat_new <- pivot_wider(exdat, id_cols=c(PPN, DOB, Gender), names_from=intervention_seq, values_from=c(Intervention_Date, Intervention_Type))`
